---
title: "The impact of covid 19 on the minimum nights of stay for airbnb's in Amsterdam"
author: "Team 1"
format: 
  revealjs:
    footer: "covid 19 on minimum nights of stay - team 1"
    theme: default
    slide-number: true
    embed-resources: true
    controls: true
    transition: slide
editor: visual
---

# Introduction

## Introduction

```{r, echo=FALSE}
# load packages
library(quarto)
library(readr)
library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
df_cleaned <- read_csv("../../gen/data-preparation/output/data_no_outliers.csv")
```

Introduction about the topic

# Motivation

## Motivation

Motivation about the topic:

# Research method

## Research method

First, we run the regression:

```{r}
#| echo: true
#| output-location: fragment
#| code-line-numbers: "2|3"
# Estimate simple model
m1 <- lm(minimum_nights ~ covid + as.factor(neighbourhood_num) + as.factor(roomtype_num) + accommodates + price + instant_bookable, df_cleaned)
summary(m1)
```

# Results

## Results {auto-animate="true"}

``` r
# Estimate simple model
summary(m1)
```
## Results {auto-animate="true"}

``` r
# Estimate simple model
m1 <- lm(minimum_nights ~ covid + as.factor(neighbourhood_num) + as.factor(roomtype_num) + accommodates + price + instant_bookable, df_cleaned)
summary(m1)
```

Type explanation about the results here

# Robustness checks

We do need to perform some robustness checks

## Independence

Type how we checked for independence

## Homodskedasticity

## Normality

## Linearity

# Conclusion

## Conclusion

Type the conclusion here
